---
title: "EDA"
output: word_document
date: "2024-05-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
### Importing important libraries
library(dplyr)
library(tidyverse)
library(ggplot2)
library(leaflet)
library(scales)
library(caret)
library(corrplot)
library(smotefamily)
library(performanceEstimation)
library(ROSE)
library(randomForest)
library(glmnet)
library(pROC)
library(e1071)
```


```{r}
# Loading data
application <- read.csv("application_record.csv")
credit<- read.csv("credit_record.csv")
```


```{r}
glimpse(credit)
```
### Target setting

STATUS: 0: 1-29 days past due 1: 30-59 days past due 2: 60-89 days overdue 3: 90-119 days overdue 4: 120-149 days overdue 5: Overdue or bad debts, write-offs for more than 150 days C: paid off that month X: No loan for the month.

Customer with over 30 days or more past due would be considered as a ‘bad’ customer. 
Based on this consideration, we can adjust the value of “STATUS” column to be 1 (bad customer) and 0 (good customer) and make it as our target for our model.

```{r}
credit$target <- credit$STATUS
unique(credit$target)

credit$target <- ifelse(credit$target == 'X' | credit$target == 'C', 0, credit$target)
credit$target <- as.integer(credit$target) 
credit$target[credit$target >= 1] <- 1 
credit_df <- aggregate(target ~ ID, data = credit, FUN = max)
unique(credit$target)

```

Target Proportion
```{r}
dplyr::count(credit_df, target, sort = TRUE)

```

### Data Cleaning
```{r}
# Changing the data type and adjusting some character values to binary
application <- application %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate_if(is.double, as.integer) %>% 
  mutate(CODE_GENDER = as.integer(ifelse(CODE_GENDER == "M", 1, 0)),
         FLAG_OWN_CAR = as.integer(ifelse(FLAG_OWN_CAR == "Y", 1, 0)),
         FLAG_OWN_REALTY = as.integer(ifelse(FLAG_OWN_REALTY == "Y", 1, 0)))
```


```{r}
#Merging credit_record and application_record
credit_data <- merge(application, credit_df, by = "ID") %>% 
  select(-c(ID))
head(credit_data)
```
### Exploratory Data Analysis

```{r}
df <- credit_data
```


```{r}
# Distribution based on gender
ggplot(df, aes(x = CODE_GENDER, fill = as.factor(target))) +
  geom_bar() +
  labs(title = "Gender Distribution by Target", x = "Gender", y = "Frequency", fill = "Target") +
  scale_fill_manual(values = c("0" = "skyblue", "1" = "red")) +
  theme_minimal()
```

Insights: 
```{r}
# Distribution based on marital status
ggplot(df, aes(x = NAME_FAMILY_STATUS, fill = as.factor(target))) +
  geom_bar() +
  labs(title = "Marital Status Distribution by Target", x = "Marital Status", y = "Frequency", fill = "Target") +
  scale_fill_manual(values = c("0" = "skyblue", "1" = "red")) +
  theme_minimal()
```


```{r}
# Distribution based on education level
ggplot(df, aes(x = NAME_EDUCATION_TYPE, fill = as.factor(target))) +
  geom_bar() +
  labs(title = "Education Level Distribution by Target", x = "Education Level", y = "Frequency", fill = "Target") +
  scale_fill_manual(values = c("0" = "skyblue", "1" = "red")) +
  theme_minimal()
```


```{r}
# Distribution based on occupation type
occupation_dist <- table(df$OCCUPATION_TYPE, df$target)
occupation_dist
```


```{r}
# 2.	How does the income level vary across different categories such as education type, family status, and housing type?

# box plot of income across different education types
ggplot(df, aes(x = NAME_EDUCATION_TYPE, y = AMT_INCOME_TOTAL, fill = NAME_EDUCATION_TYPE)) +
  geom_boxplot() +
  labs(title = "Income Distribution Across Different Education Types", x = "Education Type", y = "Income") +
  theme_minimal()
```


```{r}
## box plot of income across different family statuses
ggplot(df, aes(x = NAME_FAMILY_STATUS, y = AMT_INCOME_TOTAL, fill = NAME_FAMILY_STATUS)) +
  geom_boxplot() +
  labs(title = "Income Distribution Across Different Family Statuses", x = "Family Status", y = "Income") +
  theme_minimal()
```


```{r}
## box plot of income across different housing types
ggplot(df, aes(x = NAME_HOUSING_TYPE, y = AMT_INCOME_TOTAL, fill = NAME_HOUSING_TYPE)) +
  geom_boxplot() +
  labs(title = "Income Distribution Across Different Housing Types", x = "Housing Type", y = "Income") +
  theme_minimal()
```


```{r}
# Let's calculate the correlation between the number of children and the total annual income
cor(df$CNT_CHILDREN, df$AMT_INCOME_TOTAL)

```


```{r}
## Let's calculate age
df$AGE_YEARS <- round(-df$DAYS_BIRTH/365.24, 0)
df$YEARS_EMPLOYED <- round(-df$DAYS_EMPLOYED/365.24, 0)
df$YEARS_EMPLOYED[df$YEARS_EMPLOYED < 0] <- 0
```


```{r}
df <- subset(df, select = -c(DAYS_BIRTH, DAYS_EMPLOYED))

```


```{r}
# Age distribution of clients
ggplot(df, aes(x = AGE_YEARS)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
  labs(title = "Age Distribution of Clients", x = "Age", y = "Frequency") +
  theme_minimal()
```


```{r}
# Box plot of clients' ages by employment status
ggplot(df, aes(x = FLAG_OWN_CAR, y = AMT_INCOME_TOTAL, fill = FLAG_OWN_CAR)) +
  geom_boxplot() +
  labs(title = "Income Level Distribution by Car Ownership", x = "Car Ownership", y = "Income Level") +
  theme_minimal()
```

#7.	Can we identify any significant differences in socio-economic characteristics between clients who have defaulted on credit card payments and those who haven't?

# Perform t-test for annual income between default and non-default groups
Null Hypothesis (H0): There is no difference in the mean annual income between clients who have defaulted on credit card payments and those who haven't.

Alternative Hypothesis (H1): There is a difference in the mean annual income between clients who have defaulted on credit card payments and those who haven't.

```{r}
t_test_income <- t.test(AMT_INCOME_TOTAL ~ target, data = df)
print(t_test_income)
```
Based on the p-value and confidence interval, we can conclude that there is a statistically significant difference in annual income between clients who have defaulted on credit card payments and those who haven't. The mean income is higher for clients who have defaulted compared to those who haven't.



#### Let's perform a chi-square test to examine if there is an association between default status (e.g., defaulted on credit card payments or not) and a categorical variable ( occupation type).

```{r}
contingency_table <- table(df$NAME_FAMILY_STATUS, df$NAME_HOUSING_TYPE)
chi_sq_test_result <- chisq.test(contingency_table)
print(chi_sq_test_result)
```

Since the p-value is significantly smaller than the significance level ( 0.05), we reject the null hypothesis. This indicates that there is a statistically significant association between the "NAME_FAMILY_STATUS" and "NAME_HOUSING_TYPE" variables. In other words, there is evidence to suggest that these two categorical variables are associated with each other.


### Correlation Analysis

```{r}
numerical_columns <- sapply(df, is.numeric)
num_cols <- df[numerical_columns]
# names(num_cols)
num_cols <- num_cols[, !names(num_cols) %in% c("ID", "FLAG_MOBIL")]
correlation_matrix <- cor(num_cols)

corrplot(correlation_matrix, method = "color",
         addCoef.col = "black", tl.col = "black", tl.srt = 45)
```


```{r}
```

## Including Plots

You can also embed plots, for example:

```{r}
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
